Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-07-30 paper.tex
Added references and finished writing conclusion

    Streaming cells are commonly used in membrane science to determine the zeta potential of a membrane surface.
    This is an important factor in the design and operation of membrane based processes. \cite{Daiguji2004b}
    Streaming cells have also been investigated as a means of pumping water and generating power.\cite{Daiguji2004b, Davidson2008, Hon2004, Olthuis2005}

    In \cite{CherngHon2012}, salinity gradients are shown to produce electrical power

    \bibliographystyle{plain}
    \bibliography{library}

%\begin{thebibliography}{99}
%
%    \bibitem{CherngHon2012}
%    Kar~Cherng~Hon, Cunlu~Zhao, Chun~Yang, and Seow~Chay Low,
%    ``A method of producing electrokinetic power through forward osmosis'',
%    Appl. Phys. Lett. 101, 143902 (2012); doi: 10.1063/1.4756903
%
%    \bibitem{Daiguji2004b}
%    Hirofumi~Daiguji, Peidong~Yang, Andrew~Szeri, and Arun~Majumdar,
%    ``Electrochemomechanical Energy Conversion in Nanofluidic Channels'',
%    Nano Letters, 2004, vol. 4, 12, pp2315-2321
%
%    \bibitem{Davidson2008}
%    Christian~Davidson, Xiangchun~Xuan,
%    ``Effects of Stern layer conductance on electrokinetic energy conversion in nanofluidic channels'',
%    Electrophoresis-NEEDS-CHECKING
%
%    \bititem{Hon2012}
%    K.C.~Hon, C.~Zhao, C.~Yang, and S.C.~Low,
%    ``A method of producing electrokinetic power through forward osmosis'',
%    referenceTODO
%
%
%    \bibitem{Olthuis2005}
%    Authors,
%    ``Energy from streaming current and potential'',
%    journal
%
%    \bibitem{XuanLi2006}
%    Authors,
%    ``Thermodynamic analysis of electrkinetic energy converson'',
%    journal
%
%    \bibitem{RenStein2008}
%    Authors,
%    ``Slip-enhanced electrokinetic energy conversion in nanofluidic channels'',
%    journal
%
%    \bibitem{vanderHeyden2006}
%    Authors,
%    ``Electrokinetic energy conversion efficiency in nanofluidic channels'',
%    journal
%
%    \bibitem{DavidsonXuan2008b}
%    Authors,
%    ``Electrokinetic energy conversion in slip nanochannels'',
%    journal
%
%    \bibitem{Yang2003}
%    Authors,
%    ``Electrokinetic microchannel battery by means of electrokinetic and microfluidic phenomena'',
%    journal
%
%    \bibitem{vanderHeyden2007}
%    Authors,
%    ``Power Generation by Pressure-Driven Transport of Ions in Nanofluidic Channels'',
%    journal
%
%    \bibitem{Xie2008}
%    Authors,
%    ``Electric energy generation in single track-etched nanopores'',
%    journal
%
%    \bibitem{Daiguji2006}
%    Authors,
%    ``Theoretical studi on the efficiency of nanofluidic batteries
%
%
%\end{thebibliography}

>>>>

    Streaming cells are commonly used in membrane science to determine the zeta potential of a membrane surface.
    This is an important factor in the design and operation of membrane based processes. \cite{Daiguji2004b}
    A streaming cell contains one or more narrow channels (sometimes a porous material is used as the channel) across which a pressure gradient is applied.
    The application of a pressure gradient across the cell forces fluid through the cell and causes a streaming current, which in turn, creates voltage differential across the cell.
    This voltage differential is called the streaming potential.

    Streaming currents have been heavily investigated as a means of generating electrical energy from pressure gradients. \cite{Chang2009,Daiguji2006,Daiguji2004b,Davidson2008a,Davidson2008,CherngHon2012,Jiao2014,Lu2006,Olthuis2005,Osterle1964,Pennathur2007,Ren2008a,VanderHeyden2006,Heyden2007,Xie2008,Yang2003}
    Theoretical predictions of the efficiency of standard micro/nano-fluidic channels are 2\% for pure water and 7\% for sodium chloride. \cite{VanderHeyden2006}

    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10--16\thinspace$\mu m$.\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with 6\thinspace$\mu m$ pores. \cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a 75\thinspace$nm$ by 50\thinspace$\mu m$ silica channel. \cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a 200\thinspace$nm$ pore in an alumina membrane. \cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a 0.5\thinspace$nm$ cylindrical pore in polyethylene terephalate foil. \cite{Xie2008}
    \end{itemize}

    Techniques to induce hydrodynamic slip at the fluid-solid interface are predicted to increase this efficiency to 30-40\%. \cite{Davidson2008a, Ren2008a}
    However, experimental results utilising slip enhanced channels have not yet been reported in the literature.
    In \cite{Joly2006} it is suggested that a polarised hydrophobic surface would strongly enhance electro-osmotic effects and therefore further increase conversion efficiency.
    Hydrophobic channels have not been considered due to the exotic materials required to build them.

    We address the question of whether it is feasible to build a harvester using currently available materials.
    Such a harvester must produce enough energy to operate a microprocessor and radio communication device.
    It must collect all of its energy from a domestic water supply and be reliable.

    The remainder of this work is presented as follows.
    In section \ref{sect:streamingCell} we build and measure a simple streaming potential cell.
    We confirm the relationship between pressure and applied pressure and investigate the effect of channel height.
    In section \ref{sect:waterConsumption} we estimate a typical domestic water usage profile.
    This consumption profile will be used as the basis for estimating the amount of energy available to the harvester.
    Section \ref{sect:powerRequirements} determines the amount of energy an electronic meter would need to be useful.
    Following on this, section \ref{sect:harvesterSize} presents the required dimensions a usable harvester would need.
    And finally, we discuss the findings in the conclusion.

    \section{Readily Attainable Streaming Cell Power}
    \label{sect:streamingCell}
    \section{Estimation of Typical Household Water Consumption}
    \label{sect:waterConsumption}
    \section{Estimation of Power Requirements}
    \label{sect:powerRequirements}
    \section{Estimation of Streaming Cell Harvester}
    \label{sect:harvesterSize}
    \section{Conclusion}
    \bibliographystyle{plain}
    \bibliography{library}
